A semantics of core AUTOSAR

### 2016-04-29 paper/paper.tex
Ongoing tweaks...

Components and ports do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms. However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\Tr{R}{I}$ means runnable $R$ of component $I$ and $\Te{E}{P}{I}$ means element $E$ of port $P$ within component $I$. We use the following meta-variable conventions:

%TODO: port name convention change from U to P? (see ../prolog/semantics.eprover)
%
\[
\begin{array}{rcll}
	I	& \in		& \text{Component names} \\
	R	& \in		& \text{Runnable names} \\
	U	& \in		& \text{Port names} \\
	E	& \in		& \text{Sender-receiver element names} \\
	O	& \in		& \text{Client-server operation names} \\
	S	& \in		& \text{Inter-runnable variable names} \\
	X	& \in		& \text{Exclusive area names} \\
	\\
	A	& \in		& \text{Addresses} \\
	A	& ::=		& \Tr{R}{I}   \sep   \Te{E}{P}{I}   \sep \To{O}{P}{I}   \sep   \Ts{S}{I}   \sep   \Tx{X}{I}   \\
\end{array}
\]
%
Since AUTOSAR components can be arbitrarily nested, the component names $I$ can also be considered to have a similarly nested internal structure (component $I_1$ within component $I_2$ within component $I_3$, etc). We can fully ignore this detail here, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite{TODO2}.

The process terms of our calculus are as follows:
\begin{itemize}
\item $\runnable{\Tr{R}{I}}{TODOK}{T}{\Act}{N}$  \newline
Common state for runnable $R$ of component $I$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
\item $\rinst{\Tr{R}{I}}{C}{\Xs}{\Code}$ \newline
An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
\item $\excl{\Tx{X}{I}}{B}$ \newline
Exclusive area $X$ of component $I$, with boolean occupation state $B$.
\item $\irv{\Ts{S}{I}}{V}$ \newline
Inter-runnable variable $S$ of component $I$, currently holding value $V$.
\item $\delem{\Te{E}{U}{I}}{B}{V}$ \newline
Non-queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
\item $\qelem{\Te{E}{U}{I}}{\Vs}{TODO}$ \newline
Queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value sequence $\Vs$.
\item $\opres{\To{O}{U}{I}}{\Vs}$ \newline
Client-server operation $O$ in port $U$ of component $I$, currently holding result value sequence $\Vs$.
\item $\timer{\Tr{R}{I}}{T}$ \newline  %TODO: timer has only 2 args in the semantics so {T_p} is not described
Countdown timer state for runnable $R$ of component $I$ with $T$ seconds left of its current cycle.
\end{itemize}

@@

\item For each data element $E$ of each required port $U$ of each atomic component $I$, a term
$$
\begin{array}{c}
	\qelem{\Te{E}{U}{I}}{\epsilon}{TODO}
\end{array}
$$
if the \emph{swImplPolicy} attribute of element $E$ is \emph{queued}; otherwise, a term
$$
\begin{array}{c}
	\delem{\Te{E}{U}{I}}{\false}{\text{nodata}}
\end{array}
$$

\item For each operation $O$ of each required port $U$ of each atomic component $I$, a term
$$
\begin{array}{c}
	\opres{\Te{O}{U}{I}}{\epsilon}
\end{array}
$$

@@

\begin{gather*}
	\rinst{\adr{R:I}}{C}{\Xs}{\Enter{X}{\Cont}}  \red{\say{X:I}{\ent}}  \\
		\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Cont(\void)} \\
	\\
	\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Exit{X}{\Cont}}  \red{\say{X:I}{\ex}}  \\
		\rinst{\adr{R:I}}{C}{\Xs}{\Cont(\void)} \\
	\\
	\excl{\adr{X:I}}{\false} \red{\hear{X:I}{\ent}} \excl{\adr{X:I}}{\true} \\
	\\
	\excl{\adr{X:I}}{\true} \red{\hear{X:I}{\ex}} \excl{\adr{X:I}}{\false} \\
	\\
\end{gather*}

@@

\begin{gather*}
	\rinst{\adr{R:I}}{C}{\Xs}{\IrvWrite{X}{\Cont}}  \red{\say{X:I}{\ent}}  \\
		\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Cont(\void)} \\
	\\
	\rinst{\adr{R:I}}{C}{\adr{X:Xs}}{\Exit{X}{\Cont}}  \red{\say{X:I}{\ex}}  \\
		\rinst{\adr{R:I}}{C}{\Xs}{\Cont(\void)} \\
	\\
	\excl{\adr{X:I}}{\false} \red{\hear{X:I}{\ent}} \excl{\adr{X:I}}{\true} \\
	\\
	\excl{\adr{X:I}}{\true} \red{\hear{X:I}{\ex}} \excl{\adr{X:I}}{\false} \\
	\\
\end{gather*}

>>>>

Components and ports do not carry any dynamic state besides the state of their constituent parts, so they need no explicit representation as process terms. However, in order to still be able to talk about model elements using their locally scoped names, we employ a hierarchical addressing scheme, such that address $\Tr{R}{I}$ means runnable $R$ of component $I$ and $\Tep{E}{P}{I}$ means element $E$ of port $P$ within component $I$. We use the following meta-variable conventions:

%TODO: port name convention change from U to P? (see ../prolog/semantics.eprover)
%
\[
\begin{array}{rcll}
	I	& \in		& \text{Component names} \\
	R	& \in		& \text{Runnable names} \\
	U	& \in		& \text{Port names} \\
	E	& \in		& \text{Sender-receiver element names} \\
	O	& \in		& \text{Client-server operation names} \\
	S	& \in		& \text{Inter-runnable variable names} \\
	X	& \in		& \text{Exclusive area names} \\
	\\
	A	& \in		& \text{Addresses} \\
	A	& ::=		& \Tr{R}{I}   \sep   \Tep{E}{P}{I}   \sep \Top{O}{P}{I}   \sep   \Ts{S}{I}   \sep   \Tx{X}{I}   \\
\end{array}
\]
%
Since AUTOSAR components can be arbitrarily nested, the component names $I$ can also be considered to have a similarly nested internal structure (component $I_1$ within component $I_2$ within component $I_3$, etc). We can fully ignore this detail here, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite{TODO2}.

The process terms of our calculus are as follows:
\begin{itemize}
\item $\runnable{\Tr{R}{I}}{TODOK}{T}{\Act}{N}$  \newline
Common state for runnable $R$ of component $I$, showing $N$ current instances, $T$ seconds left until next possible instantiation, and activation state $\Act$.
\item $\rinst{\Tr{R}{I}}{C}{\Xs}{\Code}$ \newline
An instance of runnable $R$ of component $I$, currently executing $\Code$ and owning the exclusive areas $\Xs$, possibly running on behalf of client $C$.
\item $\excl{\Tx{X}{I}}{B}$ \newline
Exclusive area $X$ of component $I$, with boolean occupation state $B$.
\item $\irv{\Ts{S}{I}}{V}$ \newline
Inter-runnable variable $S$ of component $I$, currently holding value $V$.
\item $\delem{\Tep{E}{U}{I}}{B}{V}$ \newline
Non-queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value $V$ with boolean update status flag $B$.
\item $\qelem{\Tep{E}{U}{I}}{\Vs}{TODO}$ \newline
Queued sender-receiver data element $E$ in port $U$ of component $I$, currently holding value sequence $\Vs$.
\item $\opres{\Top{O}{U}{I}}{\Vs}$ \newline
Client-server operation $O$ in port $U$ of component $I$, currently holding result value sequence $\Vs$.
\item $\timer{\Tr{R}{I}}{T}$ \newline  %TODO: timer has only 2 args in the semantics so {T_p} is not described
Countdown timer state for runnable $R$ of component $I$ with $T$ seconds left of its current cycle.
\end{itemize}

@@

\item For each data element $E$ of each required port $U$ of each atomic component $I$, a term
$$
\begin{array}{c}
	\qelem{\Tep{E}{U}{I}}{\epsilon}{TODO}
\end{array}
$$
if the \emph{swImplPolicy} attribute of element $E$ is \emph{queued}; otherwise, a term
$$
\begin{array}{c}
	\delem{\Tep{E}{U}{I}}{\false}{\text{nodata}}
\end{array}
$$

\item For each operation $O$ of each required port $U$ of each atomic component $I$, a term
$$
\begin{array}{c}
	\opres{\Tep{O}{U}{I}}{\epsilon}
\end{array}
$$

@@

@@
